---
title: "data607_hw2_weizhou"
author: "Wei Zhou"
date: "February 10, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

###SQLScript to import CSV file into Mysql Database
-- loadflights.sql

DROP TABLE IF EXISTS movie;
DROP TABLE IF EXISTS friend;
DROP TABLE IF EXISTS rating;


CREATE TABLE movie (
  movie varchar(30)  NOT NULL,
  mid varchar(2) PRIMARY KEY
  );
  
LOAD DATA INFILE 'C:/ProgramData/MySQL/MySQL Server 8.0/Uploads/movie.csv' 
INTO TABLE movie 
FIELDS TERMINATED BY ',' 
ENCLOSED BY '"'
LINES TERMINATED BY '\r\n'
IGNORE 1 ROWS;

CREATE TABLE friend (
  friend char(2),
  fid varchar(2)
  );
  
LOAD DATA INFILE 'C:/ProgramData/MySQL/MySQL Server 8.0/Uploads/friend.csv' 
INTO TABLE friend
FIELDS TERMINATED BY ',' 
ENCLOSED BY '"'
LINES TERMINATED BY '\r\n'
IGNORE 1 ROWS;

CREATE TABLE rating (
mid varchar(2),
fid varchar(2),
rating integer
);

LOAD DATA INFILE 'C:/ProgramData/MySQL/MySQL Server 8.0/Uploads/rating.csv' 
INTO TABLE rating
FIELDS TERMINATED BY ',' 
ENCLOSED BY '"'
LINES TERMINATED BY '\r\n'
IGNORE 1 ROWS;


SET SQL_SAFE_UPDATES = 0;

SELECT 'movie', COUNT(*) FROM movie
  UNION
SELECT 'friend', COUNT(*) FROM friend
  UNION
SELECT 'rating', COUNT(*) FROM rating ; 


```{r echo=FALSE, results='hide',message=FALSE}
library(RMySQL)
library(dplyr)
library(ggplot2)
library(keyring)
```

### Establish Connection with MySQL Server
```{r}
keyring::key_set(service = "mysql", 
                 username = "root")

con <- dbConnect(MySQL(),    user='root', password=keyring::key_get("mysql","root"),   dbname='movies', host='localhost')
```
###Obtain data 
```{r}
rating <- dbSendQuery(con, "SELECT m.movie,f.friend,r.rating
                            FROM rating r 
                            left outer join movie m on m.mid = r.mid 
                            left outer join friend f on f.fid = r.fid
                            ")
rating_df=dbFetch(rating)
```

### Movie Names
```{r}
print(unique(rating_df$movie))
```
### Friend Names
```{r}
print(unique(rating_df$friend))
```
###Avg Rating / Movie
```{r}
movie_rating=rating_df%>%
  group_by(movie)%>%
  summarise(avg_rating = mean(rating))

ggplot(movie_rating, aes(y=avg_rating,x=movie))+geom_bar(stat = "identity")+theme(axis.text.x=element_text(angle=90, hjust=1))
  
```



