---
title: "Data607_project2_weizhou"
author: "Wei Zhou"
date: "3/10/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyr)
library(ggplot2)
```


### Read data from csv file , skip the first 15 rows
```{r}
steel <- read.csv('steel_import_1.csv',skip=15, stringsAsFactors = F, header= FALSE)[1:47,]
```

# DATA CLEAN
### Rename columns
```{r}
names(steel) <- c('Country_areas','201810_prequantity','201810_prevalue',
                                  '201809_prequantity','201809_prevalue',
                                  '201809_finalquantity','201809_finalvalue',
                                  '201710_finalquantity','201710_finalvalue',
                                  '201709_finalquantity','201709_finalvalue')
head(steel)
```

### Replace (-) with NA and convert to number,
```{r}
steel[steel=='(-)']<-NA
steel[,2:11] = steel%>%
      select(2:11)%>%
      # mutate_all(funs(str_replace(., ",", "")))%>%  
      mutate_all(as.numeric)
head(steel)
```

###  Transform dataset 
```{r}
steel_final_1 = gather(steel,year_month_type, Number,2:11, factor_key = TRUE)%>%
                separate(year_month_type, c('year_month','quant_value'),sep='_')%>%
                spread(quant_value,'Number')
head(steel_final_1,10)
```

```{r}
steel_final_2 = gather(steel,year_month_type, Number,2:11, factor_key = TRUE)%>%
                separate(year_month_type, c('year_month','quant_value'),sep='_')%>%
                spread(year_month,'Number')
head(steel_final_2,10)
```


# DATA ANALYSIS

###Questions：
#####1. Which country/area has the overall largest final quantity / value
```{r}
final_quantity = steel_final_2%>%
  filter(quant_value =='finalquantity')%>%
  rowwise()%>%
  mutate(finalquantity_overall= sum(`201709`,`201710`,`201809`,`201810`,na.rm=TRUE))%>%
  arrange(desc(finalquantity_overall))
head(final_quantity)

final_value = steel_final_2%>%
  filter(quant_value =='finalvalue')%>%
  rowwise()%>%
  mutate(finalvalue_overall= sum(`201709`,`201710`,`201809`,`201810`,na.rm=TRUE))%>%
  arrange(desc(finalvalue_overall))
head(final_value)

```
Europe has the highest overall quantity. 
```{r}
final_quantity$Country_areas <- factor(final_quantity$Country_areas, levels = final_quantity$Country_areas[order(final_quantity$finalquantity_overall)])

 g=ggplot(final_quantity,aes(y=finalquantity_overall,x=Country_areas,label=Country_areas))
 g+geom_bar(stat='identity', aes(fill=finalquantity_overall), width=.5)+
   coord_flip()
```

Europe has the highest overall value. 
```{r}
final_value$Country_areas <- factor(final_value$Country_areas, levels = final_value$Country_areas[order(final_value$finalvalue_overall)])

 g=ggplot(final_value,aes(y=finalvalue_overall,x=Country_areas,label=Country_areas))
 g+geom_bar(stat='identity', aes(fill=finalvalue_overall), width=.5)+
   coord_flip()
```


#####2. Which country has the highest unit value (Value / Quantity) 
Brazil has the hightest unit price while Singapore has the lowest unit price. 
```{r}
unit_value=data.frame(final_value$Country_areas,final_quantity$finalquantity_overall,final_value$finalvalue_overall)
names(unit_value) = c('country','quantity','value')
unit_value=unit_value%>%
  mutate(unit_value = quantity /value )
  
unit_value$country <- factor(unit_value$country, levels = unit_value$country[order(unit_value$unit_value)])

g1=ggplot(unit_value,aes(y=unit_value,x=country,label=country))
 g1+geom_bar(stat='identity', aes(fill=unit_value), width=.5)+
   coord_flip()
```
#####3. What does unit value distribution looks like?    
The unit value distribution is a normal distribution with slightly left skewed.   
The average unit value is 0.9373.

```{r}
ggplot(unit_value, aes(x=unit_value)) + 
  geom_histogram(binwidth = 0.1, color="darkblue", fill="lightblue")+
  geom_density(alpha=.2, fill="#FF6666") 
```
```{r}
summary(unit_value)
```

#####What is the defferent between preliminary and final quantity  
Most quantity difference falls in 0-100 bin. This is right skewed distribution. 
```{r}
diff  = steel_final_1%>%
  mutate(quantity_diff = finalquantity - prequantity,
         value_diff = finalvalue - prevalue)%>%
  filter(!is.na(quantity_diff))

ggplot(diff, aes(x=quantity_diff)) +
  geom_histogram(binwidth = 100, color="darkblue", fill="lightblue")+
  geom_density(alpha=.2, fill="#FF6666")
```

