---
title: "data607_hw3_weizhou"
author: "Wei Zhou"
date: "2/16/2019"
output:
  html_document:
    theme: cerulean
    highlight: pygments
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(stringr)
```

### Create data
```{r cars}
raw.data <-"555-1239Moe Szyslak(636) 555-0113Burns, C. Montgomery555-6542Rev. Timothy Lovejoy555 8904Ned Flanders636-555-3226Simpson, Homer5553642Dr. Julius Hibbert"
```

## 1. Conform to standard first_name last_name
### Extract full name
```{r}
name <- unlist(str_extract_all(raw.data, "[[:alpha:]., ]{2,}"))
name
```

### Remove Middle Name and title.
```{r}
name_1 = str_replace(name, "[A-z]{1,3}\\. ", "")
name_1
```

### Switch first last name position if there is comma in name. and remove comma
```{r}
name_2=str_replace(name_1,"(\\w+),\\s(\\w+)","\\2 \\1")
name_2
```
###Extract First name and last name
```{r}
first_last_name = data.frame(t(sapply(strsplit(name_2," "),head,2)))
names(first_last_name) = c('first','last')
first_last_name
```
##2. CWhether a character has a title
```{r}
name
title  =str_detect(name,"^[A-z]{1,3}\\. ") ### 1 to 3 alphabet positioning at the start of a string followed by '. '.
cbind(name,title)
```


## 3.whether a character has a second name 
```{r}
sec_name = str_detect(name,"\\s[A-z]\\. ") ### ' ' +second_name +'. '
cbind(name,sec_name)
```

## 4. Describe the types of string that conform to the folloing regular expressions and construct an example that is matched by he regular expression

#### 4.1  [0-9]+\\$ : one or more digit + $ sign
```{r}
regexp_4_1 = '[0-9]+\\$'
test_4_1 = 'abs1111$0110100101 '
str_extract_all(test_4_1,regexp_4_1)
```
#### 4.2 \\b[a-z]{1,4}\\b:  word edge + lower case any alphabet from a to z (frequencing from 1 time to 4 times) + word edge
```{r}
regexp_4_2 = "\\b[a-z]{1,4}\\b"
test_4_2 = 'cadfc abw'
str_extract_all(test_4_2,regexp_4_2)
```

#### 4.3 .*?\.txt$ :  . means any character   * means zero or more time  ? once or more times '\'   \. a period + txt at the end of the string
```{r}
regexp_4_3 = ".*?\\.txt$"
test_4_3 = '.\test.txt'
str_extract_all(test_4_3,regexp_4_3)
```

#### 4.4 \\d{2}/\\d{2}/\\d{4} : 2 digits / 2 digits / 4 digits
```{r}
regexp_4_4 = "\\d{2}/\\d{2}/\\d{4}"
test_4_4 = '11/11/1111'
str_extract_all(test_4_4,regexp_4_4)
```

#### 4.5 <(.+?)>.+?</\1>  

```{r}
print('<any character> any character </contents of group 1>')
regexp_4_5 = "<(.+?)>.+?</\\1>"
test_4_5 = '<span>20 flips come up heads</span>'
str_extract_all(test_4_5,regexp_4_5)
```

#### Congratulations you are a super nerd
```{r}
str_9 = "clcopCow1zmstc0d87wnkig7OvdicpNuggvhryn92Gjuwczi8hqrfpRxs5Aj5dwpn0Tanwo Uwisdij7Lj8kpf03AT5Idr3coc0bt7yczjatOaootj55t3Nj3ne6c4Sfek.r1w1YwwojigO d6vrfUrbz2.2bkAnbhzgv4R9i05zEcrop.wAgnb.SqoU65fPa1otfb7wEm24k6t3sR9zqe5 fy89n6Nd5t9kc4fE905gmc4Rgxo5nhDk!gr"
regexp_9 = "[A-Z]+"
str_extract_all(str_9, regexp_9)
```



