---
title: "Data607_hw5_weizhou"
author: "Wei Zhou"
date: "3/2/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(stringr)
library(tidyr)
library(dplyr)
```
#Read Data
```{r}
raw <- read.csv('./data.csv')
raw
```
#Data Transformation

### Remove empty line
```{r}
raw <-raw[-c(3),]
names(raw) = c('Airline','Arrival',"Los.Angeles","Phoenix","San.Diego","San.Francisco","Seattle")
raw
```

### Fill empty cell
```{r}
raw[2,'Airline'] = 'ALASKA'
raw[4,'Airline'] = 'AM WEST'
raw 
```

###Spead Arrival Status
```{r}
df <- gather(raw, City, Number, 3:7, factor_key = TRUE)%>%
       spread(Arrival,'Number')
head(df,10)
```

### Calculate delay / on time percentage
```{r}
df = df%>%
  mutate(total_number = delayed + `on time`
        , ontime_perc = `on time` / total_number
         ,delayed_perc = delayed / total_number)
df
```

#Analysis
###Compare airline overall delay percentage
```{r}
library(ggplot2)
ggplot(df, aes(x = Airline, y=delayed_perc, fill = City)) +
    geom_bar(stat="identity",position="dodge") + 
    xlab("Airlines") + ylab("Delays_percentage") 
```

```{r}
df%>%
  group_by(City,Airline)%>%
  summarise(delay_rate=sum(delayed)/sum(total_number))%>%
  spread(Airline, delay_rate)%>%
  mutate(delay_rate_diff = `AM WEST` - ALASKA)
```

According to this bar chart, we can conclude that the delay rate of AM WEST in these 5 cities are higher than ALASKA. 


###Calculate Average delay rate
```{r}
df%>%
  group_by(Airline)%>%
  summarize(avg_delay_rate = sum(delayed)/sum(total_number))%>%
  spread(Airline, avg_delay_rate)%>%
  mutate(Avg_delay_diff=`AM WEST` - ALASKA)

```
However, when comparing the the average delay rate, AM West is 2.4% lower than Alaska (13.3% v.s. 10.9%). This is because AM West's majority business is in Phoenix and it is maintaining relatively low delay rate.



